How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-24 Manuscript.tex
Updates from ShareLaTeX

The contributions of this paper are two-fold:
\be
\item
The  XTREES planner, which has not been previously pubslished. For demonstration purposes, we show XTREES
working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
can determine that some rows are preferred and some are to be avoided).  
\item An evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. With this strategy, other researchers could assess improvements or alternatives to the XTREE planner. 
\ei
The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREES. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

\section{From Prediction to Planning}

@@

@@

\section{A Frequently Asked Question}

When discussing this strategy, a common concern is as follows. XTREES is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never seen before. Therefore, it should be asked:

{\centerline {\em Can we trust the predictions made on such new examples?}}

Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
any data miner, there are two ``clouds'' of data:
\bi
\item One cloud is the space of training examples.
\item Another cloud is the space of test examples.
\ei
We should mistrust a model learned from data mining if the test examples  too far removed from the training data; i.e. if
these two clouds are too far apart. To test for ``too far'', we can run a data mining experiment to test how well
a model learned from the training data applies to the test data. Such experiments return some performance value.
Given this, two things can be said about predictions on changed examples:
\bi
\item Predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
\item More generally, predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

\begin{figure}
\includegraphics[width=1in]{_figs/2d.png} 
\caption{Training examples (in gray), test examples (in red) and 
changed exampled genreated by a planner (in green). Note how the green (displaced) cases are closer to the  training
date than the original test data
i.e. predictors trained on the training cases and
applied to the test cases should
work just as well, or better, on the displaced test cases.}\label{fig:howxy}
\end{figure}
%XXX end

That said, one thing to be said here
is that we {\em strongly} recommend that predictors are assessed prior to planning. That is
issue is explored further in the next section.

\input{tex/prediction_stats.tex}
  \section{Test Domains}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

  Note that this approach depends on having effective predictors for assessing the results.
  For the Siegmind data, this criteria was   relatively easy to achieve.
  The data in those data sets have a continuous class (runtime of the compiled system)
  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
  x264. 
  As a preliminary study, we split that data   into equal sized train:test groups
  and trained a Random Forest
  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
That is, we can be very confident that the predictors from the Siegmund data can assess
our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei
In the following, note our convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. In other work~\cite{krall14}, we are exploring tuning methods to automatically
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

>>>>

The contributions of this paper are two-fold:
\be
\item
The  XTREES planner, which has not been previously published. For demonstration purposes, we show XTREES
working to reduce software runtimes and software defects. That said, the method is quite general to any table of data with weighted class values (so that XTREES
can determine that some rows are preferred and some are to be avoided).  
\item An evaluation strategy used to comparatively assess planners. As  shown below, that strategy less us identify  inferior planners. With this strategy, other researchers could assess improvements or alternatives to the XTREE planner. 
\ei
The rest of this paper is structured as follows.
After describing our test data,
this paper  walks the reader through four methods that start
with a standard clustering algorithm, the  progresses on to XTREES. The second
half of the paper then presents an extensive evaluation of these methods using 
the Jureczko  and Siegmund data sets. This is followed by notes on related work
and validity.

\section{Preliminaries}
\subsection{From Prediction to Planning}

@@

\subsection{Trusting the Changes}\label{sect:trust}
   XTREES is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never seen before. Therefore, it must be asked:
\bi
\item
{\em Can we trust the predictions made on such new examples?}
\ei
Our answer is ``yes'', if those changes select for the {\em region of competency} of the predictor. When we assess
any data miner, there are two ``clouds'' of data:
\bi
\item One cloud is the space of training examples.
\item Another cloud is the space of test examples.
\ei
For a visualization of these clouds, see \fig{howxy}.

We should mistrust the predictions made by a  model   if it is being applied to examples far outside of its region of competency;
i.e. to test examples that are too far removed from the training data.   
To test for ``too far'', we can run a data mining experiment to test how well
a model learned from the training data applies to the test data. Such experiments return some performance value.

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as reliable as the performance value found in the data mining experiment.
Given this, one thing  can be asserted about predictions on changed examples:
\bi
\item Predictions about changes that move examples towards/away from the cloud of training data can be trusted more/less (respectively).
\ei  
Accordingly, all the methods of this paper make {\em trust-increasing} recommendations. That is, they push change examples {\em towards} the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data set. It shows: (1)~the training examples, (2)~the test examples, and (3)~the
changed  examples displaced after applying one of our planning methods.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

\begin{figure}
\includegraphics[width=2in]{_figs/twod.eps} 
\caption{Training examples (in gray), test examples (in red) and 
changed examples generated by a planner (in green). Note how the green (changed) examples are closer to the  training
date than the original test data
i.e. predictors trained on the training examples and
applied to the test examples should
work just as well, or better, on the changed test examples.}\label{fig:howxy}
\end{figure}
%XXX end

@@

That said, the above comes with two important caveats:
\bi
\item 
We   strongly recommend that predictors are assessed prior to planning. That
issue is explored further in the next section. 
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see our Validity section, later in this paper.
\ei

  \section{Test Domains}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.
  See github.com/ai-se/XTREES\#data for full access to this data.

@@

\input{tex/prediction_stats.tex}

  As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
  For the Siegmind data, this criteria was   relatively easy to achieve.
  The data in those data sets have a continuous class (runtime of the compiled system)
  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= 1 - \frac{abs(a - p)}{a}$ (and {\em higher} values are {\em better}).
This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
  x264. 
  As a preliminary study, we split that data   into equal sized train:test groups
  and trained a Random Forest
  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. This  achieved nearly perfect scores of \[s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\] 
That is, we can be very confident that the predictors from the Siegmund data can assess
our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

@@

\section{Four Planning Methods}
In his textbook on empirical methods, Cohen~\cite{cohen95} advises that any supposedly better
algorithm should be baselined against simpler alternatives. Accordingly,
this section described XTREES (which we call Method4) against three  
alternatives: 
  \bi
  \item  
  Method1, Method2 and Method3 use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	\item
 Method4 uses the decision tree learner of \fig{where}.D. 
  \ei

  Note that all our methods are {\em trust-increasing} in the manner described in \tion{trust}.

The following descriptions of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. In other work~\cite{krall14}, we are exploring tuning methods to automatically
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.
